## Wooldridge Ch.1 Computer Exercises

**C1** Use the data in WAGE1 for this exercise. (i) Find the average education level in the sample. What are the lowest and highest years of education?

```{r}
mean(wage1$educ)
min(wage1$educ)
max(wage1$educ)
```

(ii) Find the average hourly wage in the sample. Does it seem high or low?

```{r}
mean(wage1$wage)
print('Because this data set is from 1988, it seems pretty standard. ')
```

(iii) The wage data are reported in 1976 dollars. Using the Internet or a printed source, find the Consumer Price Index (CPI) for the years 1976 and 2013.

```{r}
print('CPI 1976: 56.9
      CPI 2013: 233')
```

(iv) Use the CPI values from part (iii) to find the average hourly wage in 2013 dollars. Now does the average hourly wage seem reasonable?

```{r}
mean_wage_1976 <- mean(wage1$wage)
CPI_ratio <- 233/56.9
mean_wage_1976 * CPI_ratio
print('Yes, this average hourly wage seems reasonable.')
```

(v) How many women are in the sample? How many men?

```{r}
female <- length(wage1$female[wage1$female == 1])
men <- length(wage1$female[wage1$female != 1])
paste('There are',female,'females, and',men,'men.')
```

**C2** Use the data in BWGHT to answer this question. (i) How many women are in the sample, and how many report smoking during pregnancy?

```{r}
paste(length(bwght$male[bwght$male == '0']),'females.')

```

(ii) What is the average number of cigarettes smoked per day? Is the average a good measure of the “typical” woman in this case? Explain.

```{r}
```

(iii) Among women who smoked during pregnancy, what is the average number of cigarettes smoked per day? How does this compare with your answer from part (ii), and why?

```{r}
```

(iv) Find the average of fatheduc in the sample. Why are only 1,192 observations used to compute this average?

```{r}
```

(v) Report the average family income and its standard deviation in dollars.

```{r}
```
